MSc Thesis

### 2023-05-23 manuscript/main.tex
added a table of the genes matching the peaks

\listfiles
\documentclass[12pt, letterpaper, ]{report}
\usepackage[style=authoryear, maxnames=2, maxbibnames=99]{biblatex}
\addbibresource{merged_biblio2.bib}
\addbibresource{internet_stuff.bib}
%\addbibresource{induction.bib}
%\addbibresource{gwas_silicon.bib}
\usepackage{caption}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{url}
\usepackage{lineno}

\usepackage{gensymb}
\usepackage{longtable}
\usepackage{titlesec}
\newcommand{\mycomment}[1]{}
% Change chapter heading font size to 14pt
\titleformat{\chapter}{\Large\bfseries}{\thechapter.}{20pt}{\Large}

@@

To test the prevalence of rapid silicon accumulation in canadian cereal crops, we selected three cultivars for each of oats, bread wheat, triticale, and barley. We selected cultivars on the basis of minimizing shared pedigree, and no cultivars shared more than one common ancestor within the last two crossing generations. At the start of the experiment, we germinated seeds in germination trays filled with moist sand. After four days, we transplanted germinated seedlings into 10 cm pots filled with SunGro potting mix amended with 0.17 g of silicic acid. Though potting mix and fresh water contain some amount of plant available silicon, we added the silicic acid to ensure that there would be minimal silicon limitation to the plants.We randomized the location of each pot within the growing space. A flood table bottom watered the pots with nutrient solution. We assigned each plant to one of three herbivory treatments: control, simulated herbivory, or true herbivory. We simulated herbivory by application of 1 mM methyl-jasmonate solution to the entire above-ground portion of the plant (Waterman et al. 2021b), while crickets housed in water-pik tubes provided true herbivory. Prior to introduction to the the plants, we acclimated crickets by feeding them for 72 hours on the same species used in this trial. Immediately preceding cricket application, we placed them in their tubes and starved them for 24 hours, as this increased the likelihood of the insects initiating feeding rapidly upon exposure to the test plants. Prior to harvest, we recorded whether the crickets had initiated feeding on the plants by visually inspecting the leaves for missing tissue.

@@

Recent research in inducible silicon plant defenses has focussed on the short-term dynamics of silicon uptake (\cite{waterman_short-term_2021,waterman_short-term_2021-1}). The promising results of this work has been highlighted for its potential applications in agriculture, where sensitive and rapid defensive phenotypes could improve plant performance and reduce reliance on more intensive pest-control measures. In this study, we attempted to demonstrate rapid silicifcation in four cereal crops. We failed to find evidence of rapidly induced silicon uptake in response to either methly-jasmonate application or herbivore exposure. In our study, the variation among our cultivars within species was similar in magnitude to variation among species (Figure \ref{Fig:baseline_si}), possibly obscuring the effects of our treatments. This study also differed from previous studies demonstrating rapid silicon uptake in a number of ways. The two previous studies both grew plants in liquid nutrient solution, carefully standardized to maintain consistently high silicon availability. In my study, we grew plants in potting soil amended with solid silicic acid, similar to \textcite{nascimento_silicon_2019}. In natural soil environments, the majority of plant-available silicon is derived from mineral or biogenic sources, and thus requires dissolution into the soil solution. The theoretical maximum concentration of silicic acid in soil solution is 2mM, however observed concentrations can be much below that. In our growing conditions, we applied silicic acid in excess of the average availability of phytoliths (a major source of plant-avaialbe silicon), so as to avoid soil conditions with low silicon presense. Despite this, silicon availability in the soil solution, and dissolution rates from solid to aqueous forms, may have been too low to facilitate rapid silicon uptake.

@@

For this experiment, we used a the L2 panel of Aegilops tauschii from (\cite{gaurav_population_2021}) grown at three different sites. Two of the sites were outdoors on the University of British Columbia campus, with planting occurring in the fall, while the third site was a glasshouse, where we vernalized seedlings in growth chambers prior to transplanting into the glasshouse environment. For full site details see Supplementary Table S1. Using 151 accessions, we started trays of seedlings in glasshouse or growth chamber environments. At approximately eight weeks after germination, seedlings were transplanted to their field sites. For each environment, we started four replicates of each accession. We planted the plants in a randomized block design, to minimize the effects of soil heterogeneity on our phenotype measurements. Each outdoor block was a 16 m\textsuperscript{2} square, with plants arranged $\sim$35 cm apart. Shortly after transplanting to the field sites, we applied water-soluble fertilizer to improve transplant survival, as well as slow-release fertilizer pellets. Field transplantation took place on the 15th of October 2022 and the 16th of December 2022. For the glasshouse environment, we started seedlings in growth chambers in January 2022. After 12 weeks, we moved the seedlings to vernalization chambers (4ºC, 8:16h light:dark) for eight weeks. We then transplanted these plants into 10cm square pots filled with SunGro potting mix and amended with 0.17g of silicic acid (Tixosil 68B, Solvay). Pots were arranged using the same randomized block design, adapted to fit on two flood tables. To ensure a comparable life stage across environments at time of harvest, these plants grew for three months (mid June – mid September 2022), until they had mature flower heads.

@@

To visualize the variability of silicon and manganese content within genotypes across environments, we converted elemental content into a rank, averaged ranks across environments, and plotted (Figure \ref{rank_plots}). 
To look for evidence of root exudation associated with silicon content, we tested a correlation between observed leaf silicon and manganese content. Prior to the analysis, we plotted histograms of silicon and manganese concentrations. We observed right-skew for both elements and applied a log-transformation. To control for intersite-variation we further transformed our measurements from log(concentration) to site-specific standard scores ($\frac{X - \mu}{\sigma}$). We then used the \verb|lmerTest| and \verb|MuMIn| (\cite{kuznetsova_2017_lmerTest,barton_2023_mumin}) package in R 4.2.2 (\cite{r_core_team_2022}) to implement the following model:

@@

To perform the GWAS analysis, we followed the methodology and code published in \textcite{gaurav_population_2022}. For brevity, the methodology of this manuscript only describes the steps we took using the data generated from \textcite{gaurav_population_2022}. For full details on how they generated the sequence data and prepared the final data sets refer to their manuscript. As per \textcite{gaurav_population_2022}, to reduce the computational intensity of my analysis, we prefiltered the total $k$-mer matrix to remove $k$-mers with a low chance of being informative. For each environment, we ran the GWAS, filtered $k$-mers with an association score of $<6$, and plotted the remaining $k$-mers. We calculated our bonferroni correction threshold using the \verb|knum_bonf.py| script provided by \textcite{gaurav_population_2022}.

@@

Of the approximately 1700 plants planted, [1300] produced enough leaf material for analysis. After pooling, we were left with 359 samples across three environments. Silicon content in \textit{Aegilops tuaschii} ranged from 0.784\% to 11.473\%. There were notable differences in silicon content based on the growing environment. The glasshouse plants averaged 1.450\% $\pm$ 0.032 (SE), while the two outdoor environments averaged 4.935\% $\pm$ 0.108 and 6.471\% $\pm$ 0.132. Silicon and manganese showed significant positive correlation ($\beta = 0.278 \pm 0.051, z = 5.46, p<0.0001, R^2_{marginal} = 0.05, R^2_{conditional} = 0.35 $) (Figure \ref{Fig:mn_si_regression}). Overall, my analysis revealed no regions of the \textit{Aegilops tauschii} genome that have significant associations across environments with silicon content (Figure \ref{Fig:si_peak_plot}), nor for manganese content (Figure \ref{Fig:mn_peak_plot}). For manganese, we detected a number of peaks in each environment that exceeded the threshold for significance. However, when comparing accross environments, we found no regions that consistently exceeded the threshold significance. We did however find 11 regions exceeding the significance threshold that were present in both of our growing environments with real soil.
\section{Discussion}

We observed major differences in silicon content between our growth environments, and did not find conistent correlations between genetic variation and variation in silicon content. Despite this, we did find a significant positive correlation between silicon and manganese content, supporting root exudates increasing silicon uptake.

@@

\begin{figure}[h]
        \includegraphics[scale=0.048]{images/rank_plots.png}
        \centering
        \caption{}
        \label{Fig:mn_si_regression}
\end{figure}

>>>>

\listfiles
\documentclass[12pt, letterpaper, ]{report}
\usepackage[style=authoryear, maxnames=2, maxbibnames=99, citestyle=apa]{biblatex}
\addbibresource{merged_biblio2.bib}
\addbibresource{internet_stuff.bib}
%\addbibresource{induction.bib}
%\addbibresource{gwas_silicon.bib}
\usepackage{caption}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{url}
\usepackage{lineno}
\usepackage{tabularx}
\usepackage{gensymb}
\usepackage{longtable}
\usepackage{titlesec}
\newcommand{\mycomment}[1]{}
% Change chapter heading font size to 14pt
\titleformat{\chapter}{\Large\bfseries}{\thechapter.}{20pt}{\Large}

@@

To test the prevalence of rapid silicon accumulation in canadian cereal crops, we selected three cultivars for each of oats, bread wheat, triticale, and barley. We selected cultivars on the basis of minimizing shared pedigree, and no cultivars shared more than one common ancestor within the last two crossing generations. At the start of the experiment, we germinated seeds in germination trays filled with moist sand. After four days, we transplanted germinated seedlings into 10 cm pots filled with SunGro potting mix amended with 0.17 g of silicic acid (a rate of 2.4 g silicon /kg of soil). Though potting mix and fresh water contain some amount of plant available silicon, we added the silicic acid to ensure that there would be minimal silicon limitation to the plants. We randomized the location of each pot within the growing space. A flood table bottom watered the pots with nutrient solution. We assigned each plant to one of three herbivory treatments: control, simulated herbivory, or true herbivory. We simulated herbivory by application of 1 mM methyl-jasmonate solution to the entire above-ground portion of the plant (Waterman et al. 2021b), while crickets housed in water-pik tubes provided true herbivory. Prior to introduction to the the plants, we acclimated crickets by feeding them for 72 hours on the same species used in this trial. Immediately preceding cricket application, we placed them in their tubes and starved them for 24 hours, as this increased the likelihood of the insects initiating feeding rapidly upon exposure to the test plants. Prior to harvest, we recorded whether the crickets had initiated feeding on the plants by visually inspecting the leaves for missing tissue.

@@

Recent research in inducible silicon plant defenses has focussed on the short-term dynamics of silicon uptake (\cite{waterman_short-term_2021,waterman_short-term_2021-1}). The promising results of this work has been highlighted for its potential applications in agriculture, where sensitive and rapid defensive phenotypes could improve plant performance and reduce reliance on more intensive pest-control measures. In this study, we attempted to demonstrate rapid silicification in four cereal crops. We failed to find evidence of rapidly induced silicon uptake in response to either methyl-jasmonate application or herbivore exposure.

This study differed from previous studies demonstrating rapid silicon uptake in a number of ways. The two previous studies (\cite{waterman_short-term_2021-1,waterman_short-term_2021}) both grew plants in liquid nutrient solution, carefully standardized to maintain consistently high silicon availability. In my study, we grew plants in potting soil amended with solid silicic acid, similar to \textcite{nascimento_silicon_2019}. In natural soil environments, the majority of plant-available silicon is derived from mineral or biogenic sources, and thus requires dissolution into the soil solution. The theoretical maximum concentration of silicic acid in soil solution is 2mM, however observed concentrations can be much below that. In our growing conditions, we applied silicic acid in excess of the average availability of phytoliths (a major source of plant-available silicon) (\cite{de_tombeur_plants_2020}), so as to avoid soil conditions with low silicon presence. Despite this, silicon availability in the soil solution, and dissolution rates from solid to aqueous forms, may have been too low to facilitate rapid silicon uptake.

Our observations on silicon content variation among species complement those of \textcite{simpson_still_2017}. In their study, they tested 8 species groups of wild and domesticated cereal crops, including wheat and barley. Our results for wheat and barley silicon and phenolic content more or less follow the patterns they observed, with wheat having significantly higher silicon content, and barley having much less, while the two species have more similar phenolic content. \textcite{simpson_still_2017} did not test triticale nor oats, so this study does add first records of the silicon content and phenolic content for these species. Of course, both of these traits are dependent on local environmental conditions (\cite{quigley_soil_2020,schaller_silicon_2012}), yet we are still able to interpret relative differences for plants grown under common conditions, though we likely will find quantitative variation in silicon and phenolic status among studies and growing environments. In our study, oats had the highest phenolic concentration, and the lowest silicon content. The relative difference in the phenolic and silicon content of oats and wheat in our study [was similar] to the largest difference in content in \textcite{simpson_still_2017}, but interestingly wheat is much closer to oats phylogenetically than to sorghum, the highest phenolic content species group Simpson's study. This suggests an important role for shorter term evolution in determining silicon and phenolic content.

@@

For this experiment, we used a the L2 panel of Aegilops tauschii from (\cite{gaurav_population_2021}) grown at three different sites. Two of the sites were outdoors on the University of British Columbia campus, with planting occurring in the fall, while the third site was a glasshouse, where we vernalized seedlings in growth chambers prior to transplanting into the glasshouse environment. For full site details see Supplementary Table S1. Using 151 accessions, we started trays of seedlings in glasshouse or growth chamber environments. At approximately eight weeks after germination, seedlings were transplanted to their field sites. For each environment, we started four replicates of each accession. We planted the plants in a randomized block design, to minimize the effects of soil heterogeneity on our phenotype measurements. Each outdoor block was a 16 m\textsuperscript{2} square, with plants arranged $\sim$35 cm apart. Shortly after transplanting to the field sites, we applied water-soluble fertilizer to improve transplant survival, as well as slow-release fertilizer pellets. Field transplantation took place on the 15th of October 2021 and the 16th of December 2021. For the glasshouse environment, we started seedlings in growth chambers in January 2022. After 12 weeks, we moved the seedlings to vernalization chambers (4ºC, 8:16h light:dark) for eight weeks. We then transplanted these plants into 10cm square pots filled with SunGro potting mix and amended with 0.17g of silicic acid (Tixosil 68B, Solvay). Pots were arranged using the same randomized block design, adapted to fit on two flood tables. To ensure a comparable life stage across environments at time of harvest, these plants grew for three months (mid June – mid September 2022), until they had mature flower heads.

@@

To visualize the variability of silicon and manganese content within genotypes across environments, we converted elemental content into a rank, averaged ranks across environments, and plotted (Figure \ref{Fig:rank_plots}). 
To look for evidence of root exudation associated with silicon content, we tested a correlation between observed leaf silicon and manganese content. Prior to the analysis, we plotted histograms of silicon and manganese concentrations. We observed right-skew for both elements and applied a log-transformation. To control for intersite-variation we further transformed our measurements from log(concentration) to site-specific standard scores ($\frac{X - \mu}{\sigma}$). We then used the \verb|lmerTest| and \verb|MuMIn| (\cite{kuznetsova_2017_lmerTest,barton_2023_mumin}) package in R 4.2.2 (\cite{r_core_team_2022}) to implement the following model:

@@

To perform the GWAS analysis, we followed the methodology and code published in \textcite{gaurav_population_2022}. For brevity, the methodology of this manuscript only describes the steps we took using the data generated from \textcite{gaurav_population_2022}. For full details on how they generated the sequence data and prepared the final data sets refer to their manuscript. As per \textcite{gaurav_population_2022}, to reduce the computational intensity of my analysis, we prefiltered the total $k$-mer matrix to remove $k$-mers with a low chance of being informative. For each environment, we ran the GWAS, filtered $k$-mers with an association score of $<6$, and plotted the remaining $k$-mers. We calculated our bonferroni correction threshold using the \verb|knum_bonf.py| script provided by \textcite{gaurav_population_2022}. For peaks that were significant in at least two environments, we crossreferenced the location with the Aet v5.0 assembly in NCBI (\cite{wang_aegilops_2021}), looking for annotated genes that matched our observed peak locations.

@@

 Silicon content in \textit{Aegilops tuaschii} ranged from 0.784\% to 11.473\%. There were notable differences in silicon content based on the growing environment. The glasshouse plants averaged $1.450\% \pm 0.032 (SE)$, while the two outdoor environments averaged 4.935\% $\pm$ 0.108 and 6.471\% $\pm$ 0.132. Manganese content varied from 36.5 ppm to 1296.5 ppm, and averaged $225.1 \pm 25.45 ppm (SE)$ in the glasshouse environment, and $ 71.05 \pm 2.571 ppm (SE)$ and $ 224.0 \pm 11.01 (SE)$ for the two outdoor growing environments. The results summarized in our rank plots (Figure \ref{Fig:rank_plots}) show that rank is variable for both silicon and manganese. Silicon and manganese showed significant positive correlation ($\beta = 0.278 \pm 0.051, z = 5.46, p<0.0001, R^2_{marginal} = 0.05, R^2_{conditional} = 0.35 $) (Figure \ref{Fig:mn_si_regression}). Overall, my analysis revealed no regions of the \textit{Aegilops tauschii} genome that have significant associations across environments with silicon content (Figure \ref{Fig:si_peak_plot}), nor with manganese content (Figure \ref{Fig:mn_peak_plot}). For manganese, we detected a number of peaks in each environment that exceeded the threshold for significance. However, when comparing accross environments, we found no regions that consistently exceeded the threshold significance. We did however find 11 regions exceeding the significance threshold that were present in both of our growing environments with real soil, though absent in the glasshouse environment (Table \ref{tab:peaks_and_genes}).

We observed major differences in silicon and manganese content between our growth environments, and did not find consistent correlations between genetic variation and variation in silicon content. Our rank plots revealed that relative content of both silicon and manganese is not stable across environments. This suggests strong genotype by environment interactions, likely reduced our ability to detect genetic contribution to the silicon and manganese phenotype. Interestingly, we detected a number of peaks for leaf managese content, though many of these were only found in a single environment. Though we detected no peaks present in all three growing environments, we detected 11 regions that were significant in both outdoor environments. Compared to the glasshouse environment, the outdoor environments likely exposed the plants to a soil environment in which root exudation yeilds nutrients for uptake. Root exudation of organic acids and chelating agents liberates phosphates and silicates from inorganic pools (\cite{de_tombeur_silicon_2021-1,lambers_plant_2008}). In the glasshouse environment, we grew plants in standard potting mix, which is high in organic matter with a relatively low inorganic component. In this soil context, root exudation is unlikely to yeild appreciable nutrient liberation. The high manganese content found in the glasshouse plants is most likely due to the watering regime they experienced. We watered these plants using municipal tap water. In the Vancouver region, this water is sourced from reservoirs in the surrounding mountains, which themselves are filled from ground runoff. Nearby groundwater wells have issues with high manganese content (\cite{hu_drinking_2020}). A constant supply of soluble manganese in the irrigation the glasshouse plants likely accounts for the elevated manganese content.

Our observations of a correlation between silicon and manganese content lend support to the idea of root-exudate based silicon aquisition (\cite{de_tombeur_silicon_2021-1}). It is possible that this positive correlation is due to a nutrient effect more generally  
\section{Acknowledgements}

@@

\begin{table}[ht]
        \centering
        \caption{GWAS peaks for manganese content crossreferenced with the \textit{Aegilops tauschii} Aet v5.0 assembly. }
        \label{tab:peaks_and_genes}
        \begin{tabular}{|lrrr|}
        \hline
        \textbf{Chromosome} & \textbf{Peak Location}                & \textbf{Range}                             & \textbf{Gene Symbol} \\
        \hline
        1D         & 201950k-201960k   & No Gene Found                     &                  \\
        1D         & 402540k-402550k   & 402549667-402554765              & LOC109755172   \\
        1D         & 402540k-402550k   & 402547029-402549598 & LOC120968250      \\
        1D         & 418870k-418880k   & No Gene Found                     &                  \\
        1D         & 418990k-419000k   & No Gene Found                     &                  \\
        1D         & 419030k-419040k   & No Gene Found                     &                  \\
        3D         & 575420k-575430k   & 575424894-575428979 & LOC109751197   \\
        4D         & 85300k-85310k     & No Gene Found                     &                  \\
        4D         & 85310k-85320k     & No Gene Found                     &                  \\
        4D         & 85330k-85340k     & No Gene Found                     &                  \\
        \hline
        \end{tabular}
\end{table}

\begin{figure}[h]
        \includegraphics[scale=0.048]{images/rank_plots.png}
        \centering
        \caption{Average rank of silicon and manganese content of 113 genotypes of \textit{Aegilops tauschii} grown in three different environments. Error bars represent the standard deviation of rank.}
        \label{Fig:rank_plots}
\end{figure}
